\thetitle

### 2013-04-17 tfpie/tfpie.tex
more on profj
Use tree example.

\subsection{Dispatch}

Unions, and recursive unions in particular, are a fundamental kind of
data definition that students are well-versed in from the previous
semester.  A fundamental early lesson is how to represent (recursive)
unions using classes and how to write recursive methods.  As an
example, here is a simple class definition for lists of numbers
(an archetypical recursive union data definition) that includes
a method for computing the length of a list:
\begin{verbatim}
    ;; A list of numbers (LoN) is one of:
    ;; - (new empty%)
    ;; - (new cons% Number LoN)

    (define-class empty%
      ;; Length of this empty list
      (check-expect ((new empty%) . len) 0)
      (define (len) 0))

    (define-class cons% (fields first rest)
      ;; Length of this cons list
      (check-expect ((new cons% 7 (new empty%)) . len) 1)
      (define (len)
        (add1 (this . rest . len))))
\end{verbatim}

The structure of this data definition is analagous to the approach of
the previous semester but this example brings to light an important
difference with the functional approach.

\subsection{Day 1}

[First semester vs second semester example programs.]

In intermediate student:

@@

(check-expect ((new leaf% 7) . sum) 7)
(check-expect 
  ((new node%
        (new leaf% 1)
        5 
        (new node% (new leaf% 0) 
                   10
                   (new leaf% 0))))
    . sum)
  16)
\end{verbatim}
\end{minipage}
\caption{Binary tree sum in Beginning Student and in our language}
\end{figure}

@@

However, the ``day 1'' transition from the student languages used with
\htdp to \profj is too abrupt and too large.
%
Most significantly, changing languages from the first semester
immediately rather than simply adding a new concept confuses too many
issues for students.  On the first day of a \htdc-based course,
students see object-orientation, a new programming paridigm; Java, a
new language with new syntax, and static types, a crucial but
orthogonal concept.

>>>>

\subsection{Where did the {\tt cond} go?}

Unions, and recursive unions in particular, are a fundamental kind of
data definition that students are well-versed in from the previous
semester.  A fundamental early lesson is how to represent (recursive)
unions using classes and how to write recursive methods.  As an
example, figure~\ref{fig:tree} defines binary trees of numbers
(an archetypical recursive union data definition)
using the BSL language and the Class language.
%% \begin{verbatim}
%%     ;; A list of numbers (LoN) is one of:
%%     ;; - (new empty%)
%%     ;; - (new cons% Number LoN)

%%     (define-class empty%
%%       ;; Length of this empty list
%%       (check-expect ((new empty%) . len) 0)
%%       (define (len) 0))

%%     (define-class cons% (fields first rest)
%%       ;; Length of this cons list
%%       (check-expect ((new cons% 7 (new empty%)) . len) 1)
%%       (define (len)
%%         (add1 (this . rest . len))))
%% \end{verbatim}

@@

(check-expect ((new leaf% 7) . sum) 7)
(check-expect 
  ((new node%
        (new leaf% 1)
        5 
        (new node% (new leaf% 0) 
                   10
                   (new leaf% 0))))
    . sum)
  16)
\end{verbatim}
\end{minipage}
\caption{Binary tree sum in Beginning Student and in the Class language}
\label{fig:tree}
\end{figure}

The structure of this data definition is analagous to the approach of
the previous semester but this example brings to light an important
difference with the functional approach.  The method for computing the
sum of the empty tree is defined in the {\tt leaf\%} class, while the
method for computing the sum of a node is in the {\tt node\%} class.
When a tree object is sent the {\tt sum} method, there is no function
with a conditional to determine whether the object is a
leaf---instead, the object itself takes care of computing the sum
based on the {\tt sum} method it contains.  This shift in perspective
is at the core of object-orientation: objects contain their own
behavior and the case analysis previously done in functions is
eliminated.

@@

However, the ``day 1'' transition from the student languages used with
\htdp to \profj is too abrupt and too large.
%
Most significantly, changing languages from the first semester
immediately rather than simply adding a new concept confuses too many
issues for students.  On the first day of a \htdc-based course,
students see object-orientation, a new programming paridigm; Java, a
new language with new syntax, and static types, a crucial but
orthogonal concept.
%
In contrast, our course presents just one of these concepts on the
first day, but  covers all of them by the end of the semester.
